---
title: "Continuous Mixture Distributions"
author: "Allison C Fialkowski"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: Bibliography.bib
vignette: >
  %\VignetteIndexEntry{Continuous Mixture Distributions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, fig.width = 6, fig.height = 4)
```

The following examples demonstrate the use of the `contmixvar1` function, which simulates one continuous mixture distribution $Y$.  @Head2002's fifth-order transformation is used to generate the components of $Y$ with `n = 10000`.  @HeadKow outlined a general method for comparing a simulated distribution $Y$ to a given theoretical distribution $Y^*$.  These have been modified below to apply to continuous mixture variables and will be shown for **Example 1**.  Note that they could easily be modified further for comparison to an empirical vector of data.  More details about continuous mixture variables can be found in the [Expected Cumulants and Correlations for Continuous Mixture Variables](corr_mixture.html) vignette.  Some code has been modified from the **SimMultiCorrData** package [@SMCD].

1. **Obtain the standardized cumulants** (skewness, kurtosis, fifth, and sixth) for the components of $Y^*$.  This can be done using `SimMultiCorrData::calc_theory` along with either the component distribution names (plus up to 4 parameters for each) or the PDF's with support bounds.  In the case of an empirical vector of data, use `SimMultiCorrData::calc_moments` or `SimMultiCorrData::calc_fisherk` [@SMCD].  If you desire $Y$ to have the mean and variance of $Y^*$, use `calc_mixmoments` with the component parameters as inputs to find $E[Y^*]$ and $SD[Y^*]$.

1. **Obtain the constants** for the components of $Y$.  This can be done using `SimMultiCorrData::find_constants` on each component or by simulating the mixture variable with `contmixvar1`.

1. Determine whether these constants produce a **valid power method PDF**.  The results of `find_constants` or `contmixvar1` indicate whether the constants yield invalid or valid PDF's for the component variables.  The constants may also be checked using `SimMultiCorrData::pdf_check`.  If the constants generate an invalid PDF, the user should check if the skurtosis falls above the lower bound (using `SimMultiCorrData::calc_lower_skurt`).  If yes, sixth cumulant correction values should be used in `SimMultiCorrData::find_constants` or `contmixvar1` to find the smallest corrections that produce valid PDF constants.  If all of the component distributions have valid PDF's, the mixture variable has a valid PDF.

1. **Select a critical value** from $Y^*$, i.e. $y^*$ such that $Pr(Y^* \ge y^*) = \alpha$ for the desired significance level $\alpha$.

1. **Calculate** the cumulative probability for the simulated mixture variable $Y$ up to $y^*$ and compare to $1 - \alpha$.

1. **Plot a parametric graph** of $Y^*$ and $Y$.  This can be done with the simulated vector of data $Y$ using `plot_simpdf_theory` (`overlay` = TRUE) and specifying the PDF `fx` for the mixture variable.  If comparing to an empirical vector of data, use `SimMultiCorrData::plot_sim_pdf_ext`.

```{r, include=FALSE}
library("SimCorrMix")
library("printr")
```

## Example 1: Mixture of 2 Normal Distributions

The component distributions are *Normal(-2, 1)* and *Normal(2, 1)*.  The mixing proportions are $0.4$ and $0.6$.  Use @HeadKow's steps to compare the density of the simulated variable to the theoretical density:

### Step 1: Obtain the standardized cumulants

The values of $\gamma_1,\ \gamma_2,\ \gamma_3,$ and $\gamma_4$ are all $0$ for normal variables.  The mean and standard deviation of the mixture variable are found with `calc_mixmoments`.
```{r}
options(scipen = 999)
n <- 10000
mix_pis <- c(0.4, 0.6)
mix_mus <- c(-2, 2)
mix_sigmas <- c(1, 1)
mix_skews <- rep(0, 2)
mix_skurts <- rep(0, 2)
mix_fifths <- rep(0, 2)
mix_sixths <- rep(0, 2)
Nstcum <- calc_mixmoments(mix_pis, mix_mus, mix_sigmas, mix_skews, 
  mix_skurts, mix_fifths, mix_sixths)
```

### Step 2: Simulate the variable

Note that `calc_mixmoments` returns the standard deviation, not the variance.  The simulation functions require variance as the input.  First, the parameter inputs are checked with `validpar`.
```{r}
validpar(k_mix = 1, method = "Polynomial", means = Nstcum[1], 
  vars = Nstcum[2]^2, mix_pis = mix_pis, mix_mus = mix_mus, 
  mix_sigmas = mix_sigmas, mix_skews = mix_skews, mix_skurts = mix_skurts, 
  mix_fifths = mix_fifths, mix_sixths = mix_sixths)
Nmix2 <- contmixvar1(n, "Polynomial", Nstcum[1], Nstcum[2]^2, mix_pis, mix_mus, 
  mix_sigmas, mix_skews, mix_skurts, mix_fifths, mix_sixths)
```

Look at a summary of the target distribution and compare to a summary of the simulated distribution.
```{r}
knitr::kable(Nmix2$target_mix, digits = 5, row.names = FALSE, 
  caption = "Summary of Target Distribution")
knitr::kable(Nmix2$mix_sum, digits = 5, row.names = FALSE, 
  caption = "Summary of Simulated Distribution")
```

### Step 3: Determine if the constants generate a valid power method PDF

```{r}
Nmix2$constants
Nmix2$valid.pdf
```

### Step 4: Select a critical value

Let $\alpha = 0.05$.  Since there are no quantile functions for mixture distributions, determine where the cumulative probability equals $1 - \alpha = 0.95$.  The boundaries for `uniroot` were determined through trial and error.
```{r}
fx <- function(x) 0.4 * dnorm(x, -2, 1) + 0.6 * dnorm(x, 2, 1)
cfx <- function(x, alpha, FUN = fx) {
  integrate(function(x, FUN = fx) FUN(x), -Inf, x, subdivisions = 1000, 
    stop.on.error = FALSE)$value - (1 - alpha)
}
y_star <- uniroot(cfx, c(3.3, 3.4), tol = 0.001, alpha = 0.05)$root
y_star
```

### Step 5: Calculate the cumulative probability for the simulated variable up to $1 - \alpha$

We will use the function `SimMultiCorrData::sim_cdf_prob` to determine the cumulative probability for $Y$ up to `y_star`.  This function is based on Martin Maechler's `ecdf` function [@Stats].
```{r}
sim_cdf_prob(sim_y = Nmix2$Y_mix[, 1], delta = y_star)$cumulative_prob
```

This is approximately equal to the $1 - \alpha$ value of $0.95$, indicating the method provides a **good approximation to the actual distribution.**

### Step 6: Plot graphs

```{r}
plot_simpdf_theory(sim_y = Nmix2$Y_mix[, 1], ylower = -10, yupper = 10, 
  title = "PDF of Mixture of Normal Distributions", fx = fx, lower = -Inf, 
  upper = Inf)
```

We can also plot the empirical cdf and show the cumulative probability up to y_star.
```{r}
plot_sim_cdf(sim_y = Nmix2$Y_mix[, 1], calc_cprob = TRUE, delta = y_star)
```

## Example 2: Mixture of 3 Beta Distributions

The component distributions are *Beta(6, 3)*, *Beta(4, 1.5)*, and *Beta(10, 20)*.  The mixing proportions are $0.5, 0.2, 0.3$.

```{r}
Stcum1 <- calc_theory("Beta", c(6, 3))
Stcum2 <- calc_theory("Beta", c(4, 1.5))
Stcum3 <- calc_theory("Beta", c(10, 20))
mix_pis <- c(0.5, 0.2, 0.3)
mix_mus <- c(Stcum1[1], Stcum2[1], Stcum3[1])
mix_sigmas <- c(Stcum1[2], Stcum2[2], Stcum3[2])
mix_skews <- c(Stcum1[3], Stcum2[3], Stcum3[3])
mix_skurts <- c(Stcum1[4], Stcum2[4], Stcum3[4])
mix_fifths <- c(Stcum1[5], Stcum2[5], Stcum3[5])
mix_sixths <- c(Stcum1[6], Stcum2[6], Stcum3[6])
mix_Six <- list(seq(0.01, 10, 0.01), c(0.01, 0.02, 0.03), seq(0.01, 10, 0.01))
Bstcum <- calc_mixmoments(mix_pis, mix_mus, mix_sigmas, mix_skews, mix_skurts, 
  mix_fifths, mix_sixths)
validpar(k_mix = 1, method = "Polynomial", means = Bstcum[1], 
  vars = Bstcum[2]^2, mix_pis = mix_pis, mix_mus = mix_mus, 
  mix_sigmas = mix_sigmas, mix_skews = mix_skews, mix_skurts = mix_skurts, 
  mix_fifths = mix_fifths, mix_sixths = mix_sixths, mix_Six = mix_Six)
Bmix3 <- contmixvar1(n, "Polynomial", Bstcum[1], Bstcum[2]^2, mix_pis, mix_mus, 
  mix_sigmas, mix_skews, mix_skurts, mix_fifths, mix_sixths, mix_Six)
knitr::kable(Bmix3$target_mix, digits = 5, row.names = FALSE, 
  caption = "Summary of Target Distribution")
knitr::kable(Bmix3$mix_sum, digits = 5, row.names = FALSE, 
  caption = "Summary of Simulated Distribution")
```

Is this a valid PDF?
```{r}
Bmix3$valid.pdf
Bmix3$sixth_correction
Bmix3$constants
Bplot <- plot_simpdf_theory(Bmix3$Y_mix[, 1], 
  title = "PDF of Mixture of Beta Distributions",
  fx = function(x) mix_pis[1] * dbeta(x, 6, 3) + 
    mix_pis[2] * dbeta(x, 4, 1.5) + mix_pis[3] * dbeta(x, 10, 20), lower = 0, 
  upper = 1)
Bplot
```

## Example 3: Mixture of 2 Chi-square Distributions

The component distributions are *Chisq(2)* and *Chisq(32)*.  The mixing proportions are $0.7$ and $0.3$.

```{r}
Stcum1 <- calc_theory("Chisq", 2)
Stcum2 <- calc_theory("Chisq", 32)
mix_pis <- c(0.7, 0.3)
mix_mus <- c(Stcum1[1], Stcum2[1])
mix_sigmas <- c(Stcum1[2], Stcum2[2])
mix_skews <- c(Stcum1[3], Stcum2[3])
mix_skurts <- c(Stcum1[4], Stcum2[4])
mix_fifths <- c(Stcum1[5], Stcum2[5])
mix_sixths <- c(Stcum1[6], Stcum2[6])
Cstcum <- calc_mixmoments(mix_pis, mix_mus, mix_sigmas, mix_skews, mix_skurts, 
  mix_fifths, mix_sixths)
validpar(k_mix = 1, method = "Polynomial", means = Cstcum[1], 
  vars = Cstcum[2]^2, mix_pis = mix_pis, mix_mus = mix_mus, 
  mix_sigmas = mix_sigmas, mix_skews = mix_skews, mix_skurts = mix_skurts, 
  mix_fifths = mix_fifths, mix_sixths = mix_sixths)
Cmix2 <- contmixvar1(n, "Polynomial", Cstcum[1], Cstcum[2]^2, mix_pis, mix_mus, 
  mix_sigmas, mix_skews, mix_skurts, mix_fifths, mix_sixths)
knitr::kable(Cmix2$target_mix, digits = 5, row.names = FALSE, 
  caption = "Summary of Target Distribution")
knitr::kable(Cmix2$mix_sum, digits = 5, row.names = FALSE, 
  caption = "Summary of Simulated Distribution")
```

Is this a valid PDF?
```{r}
Cmix2$valid.pdf
Cmix2$constants
Cplot <- plot_simpdf_theory(Cmix2$Y_mix[, 1], 
  title = "PDF of Mixture of Chi-square Distributions",
  fx = function(x) mix_pis[1] * dchisq(x, 2) + mix_pis[2] * dchisq(x, 32), 
  lower = 0, upper = Inf)
Cplot
```

## Example 4: Mixture of 4 Normal Distributions

The component distributions are *Normal(-4, 1)*, *Normal(-2, 2)*, *Normal(1, 3)*, and *Normal(5, 4)*.  The mixing proportions are $0.3$, $0.2$, $0.4$, and $0.1$.

```{r}
mix_pis <- c(0.3, 0.2, 0.4, 0.1)
mix_mus <- c(-4, -2, 1, 5)
mix_sigmas <- c(1, 2, 3, 4)
mix_skews <- rep(0, 4)
mix_skurts <- rep(0, 4)
mix_fifths <- rep(0, 4)
mix_sixths <- rep(0, 4)
Nstcum <- calc_mixmoments(mix_pis, mix_mus, mix_sigmas, mix_skews, mix_skurts, 
  mix_fifths, mix_sixths)
validpar(k_mix = 1, method = "Polynomial", means = Nstcum[1], 
  vars = Nstcum[2]^2, mix_pis = mix_pis, mix_mus = mix_mus, 
  mix_sigmas = mix_sigmas, mix_skews = mix_skews, mix_skurts = mix_skurts, 
  mix_fifths = mix_fifths, mix_sixths = mix_sixths)
Nmix4 <- contmixvar1(n, "Polynomial", Nstcum[1], Nstcum[2]^2, mix_pis, mix_mus, 
  mix_sigmas, mix_skews, mix_skurts, mix_fifths, mix_sixths)
knitr::kable(Nmix4$target_mix, digits = 5, row.names = FALSE, 
  caption = "Summary of Target Distribution")
knitr::kable(Nmix4$mix_sum, digits = 5, row.names = FALSE, 
  caption = "Summary of Simulated Distribution")
```

Is this a valid PDF?
```{r}
Nmix4$valid.pdf
Nmix4$constants
Nplot <- plot_simpdf_theory(Nmix4$Y_mix[, 1], 
  title = "PDF of Mixture of Normal Distributions",
  fx = function(x) mix_pis[1] * dnorm(x, mix_mus[1], mix_sigmas[1]) + 
    mix_pis[2] * dnorm(x, mix_mus[2], mix_sigmas[2]) + 
    mix_pis[3] * dnorm(x, mix_mus[3], mix_sigmas[3]) + 
    mix_pis[4] * dnorm(x, mix_mus[4], mix_sigmas[4]), 
  lower = -Inf, upper = Inf)
Nplot
```

## References

<script type="text/x-mathjax-config">
   MathJax.Hub.Config({  "HTML-CSS": { minScaleAdjust: 115, availableFonts: [] }  });
</script>
